Tracking $p$-adic precision

### 2014-02-24 padicprec.tex
Changes from Monday meeting

\section{Problems in Precision} \label{sec:prec-proposal}

We begin by offering two examples that illustrate key features of $p$-adic precision.
Recall that an element $a + O(p^N)$ of $\Qp$ has \emph{absolute precision} $N$ if
it is well defined up to adding an unknown element of $p^N\Zp$.

@@

\subsection{Approximate elements}

@@

Note first that using an arbitrary lattice to represent the precision of an approximate element
can reduce precision loss when compared to storing the precision of each coefficient
$\alpha_i$ in \eqref{eq:exact_elt} separately.  Consider the function 
$f : \Qp^2 \to \Qp^2$ mapping $(x,y)$ to $(x+y, x-y)$ and the problem of computing
$f \circ f(a + O(p^n), b + O(p^m))$.

Write $(e_1, e_2)$ for the standard basis of $E=\Qp^2$.  Since $f$ is linear, the image of
the approximation $\bigl((a,b), B_E\bigl((e_1,e_2),(p^{-n}, p^{-m})\bigr)\bigr)$ is
$\bigl((a+b, a-b), B_E\bigl((e_1+e_2, e_1-e_2), (p^{-n}, p^{-m})\bigr)\bigr)$.
For $p \ne 2$, applying $f$ again yields $\bigl((2a, 2b), B_E\bigl((e_1, e_2), (p^{-n}, p^{-m})\bigr)\bigr)$.
Storing precisions on each coordinate of a vector amounts to only allowing precisions of the form
$B_E((r_i))$ with respect to the standard basis.  One would be forced to truncate the intermediate precision
to $B_E(p^{-\min(m, n)})$, yielding an artificial loss of precision in the final result.

@@

In addition to allowing for more efficient computation of precision, storing the precision separately from the
approximation makes it easier to determine the precision of the result when step-by-step precision tracking
does not reflect the actual precision of the result.

\section{The main Lemma}
\label{sec:mainlemma}

@@

\section{Tracking precision using differentials}
\label{sec:tracking}

@@

A final benefit of separating precision from approximation is that it makes
it much easier to implement algorithms capable of processing different precision types.

>>>>

\section{Precision proposals} \label{sec:prec-proposal}

\subsection{Problems in precision}

We begin by offering two examples that illustrate key features of $p$-adic precision.
Recall that an element $a + O(p^N)$ of $\Qp$ has \emph{absolute precision} $N$ if
it is well defined up to adding an unknown element of $p^N\Zp$.
The following problems illustrate cases where step-by-step precision tracking
does not yield optimal results.

\subsubsection*{A linear map.}

Consider the function $f : \Qp^2 \to \Qp^2$ mapping $(x,y)$ to $(x+y, x-y)$ and the problem
of computing $f \circ f(a + O(p^n), b + O(p^m))$.  Applying $f$ twice, computing precision
with each step, yields $\left(2a + O(p^{\min(m,n)}), 2b + O(p^{\min(m, n)})\right)$.
On the other hand, $f \circ f(x,y) = (2x, 2y)$, so one may compute the result more
accurately as $(2a + O(p^n), 2b + O(p^m))$, with even more precision when $p=2$.

@@

\subsubsection*{Approximate elements.}

@@

Note first that using an arbitrary lattice to represent the precision of an approximate element
can reduce precision loss when compared to storing the precision of each coefficient
$\alpha_i$ in \eqref{eq:exact_elt} separately.  Recall the map 
$f : (x,y) \mapsto (x+y, x-y)$ from the beginning of the section, and write $(e_1, e_2)$
for the standard basis of $E=\Qp^2$.  Since $f$ is linear, the image of
the approximation $\bigl((a,b), B_E\bigl((e_1,e_2),(p^{-n}, p^{-m})\bigr)\bigr)$ is
$\bigl((a+b, a-b), B_E\bigl((e_1+e_2, e_1-e_2), (p^{-n}, p^{-m})\bigr)\bigr)$.
For $p \ne 2$, applying $f$ again yields $\bigl((2a, 2b), B_E\bigl((e_1, e_2), (p^{-n}, p^{-m})\bigr)\bigr)$.
By using lattices one eliminates the loss of precision seen previously.

@@

\section{Lattices and differentials}
\label{sec:mainlemma}

@@

\section{Precision in practice}
\label{sec:tracking}

@@

Note that sometimes the precision of a final result can be computed
a priori (using the methods of Section \ref{sec:differentials} for example).
Taking advantage of such knowledge can minimize artificial precision loss
even when using rougher precision types such as flat or jagged.
Separating precision from approximation also makes
it much easier to implement algorithms capable of processing different precision types,
since one can implement the arithmetic of the approximation separately from the logic
handling the precision.
